---
title: "59 - Skittles From a Bag"
author: "Felipe Leandro Aguazaco"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: darkly
    highlight: zenburn
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      cache = TRUE,
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      dev = "png")
```

# Problem

---

A bag contains 40 skittles: <span style = "color:yellow">16 yellow</span>, <span style = "color:red">14 red</span>, and <span style = "color:orange">10 orange</span>. You draw **3** skittles at random (without replacement) from the bag. What is the probability that you get 2 skittles of one color and another skittle of a different color? 

## Data 

---

```{r}

# A bag contains 40 skittles: 16 yellow, 14 red, and 10 orange.
skittles <- c(rep("Yellow", 16), rep("Red", 14), rep("Orange", 10))

# Skittles amount
length(skittles) 
skittles

```

## Example Result

---

Single experiment of draw **3** skittles at random (without replacement) from the bag:

```{r}

set.seed(20211222) # For reproducibility
draw_one <- sample(x = skittles, 
                   size = 3, 
                   replace = FALSE)

draw_one
```

Possible outputs from each experiment:

1. Three skittles of same color: Red - Red - Red  
2. Two skittles of one color and another skittle of a different color: Red - Yellow - Yellow  
3. The three skittles show different color: Red - Yellow - Orange  

# Solution

---

Two approaches:

* Monte Carlo Simulation 
* Probability and Counting Theory

## Monte Carlo Simulation 

---

> Repeat the experiment a large enough number of times, equivalent to doing it over and over. Stochastic process. 

```{r}
# 1. Experiment
draw_one

# 2. Duplicate elements
duplicated(draw_one)

# 3. If the sum of logical duplicate elements is 1, then there are two skittles of one color and another skittle of a different color
sum(duplicated(draw_one))

```

Using the `replicate` function to repeat the experiment a million times:

```{r }

library(dplyr)  # if_else function

results <- replicate(n = 1000000, {
  # 1. draw 3 skittles at random
  draws <- sample(x = skittles, 
                  size = 3, 
                  replace = FALSE)
  # 2. 
  sum <- sum(duplicated(draws))
  # 3. If the sum is different to 1 then takes a value zero
  experiment <- if_else(sum == 1, 1, 0)
})
```

The mean of the output vector *results* is the probability of getting 2 skittles of one color and another skittle of a different color:

```{r}

mean(results)

```

The probability that you get 2 skittles of one color and another skittle of a different color from a bag that contains 40 skittles: <span style = "color:yellow">16 yellow</span>, <span style = "color:red">14 red</span>, and <span style = "color:orange">10 orange</span> is **`r round(mean(results), 3)`** or **`r round(mean(results), 3)*100`**\%

## Probability and Counting Theory

---

There are three ways to select two colors (a, b) given the condition one is repeated: 

* (a, a, b) 
* (a, b, a) 
* (b, a, a)

```{r}

# Two yellow skittles and another of different color
((16/40)*(15/39)*(24/38))*3 + 
# Two red skittles and another of different color
((14/40)*(13/39)*(26/38))*3 + 
# Two orange  skittles and another of different color
((10/40)*(9/39)*(30/38))*3 

```

