---
title: "60 - Analyzing Employee Data"
author: "Felipe Leandro Aguazaco"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: darkly
    highlight: zenburn
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      dev = "png")
```

# Problem

---

Below is a snippet from a table that contains information about employees that work at Company XYZ: 

| Column name       | Data type | Example value | Description                                           |
|-------------------|-----------|---------------|-------------------------------------------------------|
| employee_name     | string    | Cindy         | Name of employee                                      |
| employee_id       | integer   | 1837204       | Unique id for each employee                           |
| yrs_of_experience | integer   | 14            | total working years of experience                     |
| yrs_at_company    | integer   | 10            | total working years at Company XYZ                    |
| compensation      | integer   | 100000        | dollar value of employee compensation                 |
| career_track      | string    | technical     | Potential values: technical, non-technical, executive |

Company XYZ Human Resource department is trying to understand compensation across the company and asked you to pull data to help them make a decision regarding employee compensation.   

Can you pull the average, median, minimum, maximum, and standard deviations for salary across 5 year experience buckets at Company XYZ? (e.g. get the corresponding average, median, minimum, maximum, and standard deviations for experience buckets 0-5, 5-10, 10-15, etc.) You can assume the data is imported into a dataframe named, df.  

# Resolving Problem {.tabset .tabset-fade .tabset-pills}

---

## In Python 

Libraries:

```{python}
import pandas as pd
import numpy as np

```

Example data:

```{python}

df = pd.DataFrame({'employee_name': ["Cindy", "Leandro", "Cristian", "Felipe", "Marina", "Luz", "Daniel", "Joseph", "Charles", "Tom"], 
                   'employee_id': list(range(1, 11, 1)), 
                   'yrs_of_experience': [4, 3, 5, 7, 8, 6, 10, 11, 20, 21], 
                   'yrs_at_company': [1, 2, 2, 1, 4, 4, 3, 3, 5, 6], 
                   'compensation': [100000, 100000, 150000, 150000, 230000, 134000, 456000, 321000, 123000, 123000], 
                   'career_track': ["technical" for i in range(1, 5)] + ["non-technical" for i in range(1, 5)] + ["executive" for i in range(1, 3)]})

df.info()
df.describe()

```

Discretize experience variable using `cut` function from `pandas` library:

```{python}
df['interval'] = pd.cut(x = df.yrs_of_experience, 
                        bins = range(0, 30, 5), # Interval limits
                        right = True)
```

Group by and aggregate functions using `pandas`:

```{python}
df.groupby(by = "interval") \
  .agg(func = ["count", "mean", "median", "min", "max", "std"]) \
  .round(decimals = 2)["compensation"]

```

## In R

Libraries:

```{r}
library(tidyverse)
library(reticulate)
```

Creating data frame in R:

```{r}

df_in_R <- data.frame(employee_name =  c("Cindy", "Leandro", "Cristian", "Felipe", "Marina", "Luz", "Daniel", "Joseph", "Charles", "Tom"),
                      employee_id= seq(1, 10, 1), 
                      yrs_of_experience = c(4, 3, 5, 7, 8, 6, 10, 11, 20, 21),
                      yrs_at_company = c(1, 2, 2, 1, 4, 4, 3, 3, 5, 6), 
                      compensation = c(100000, 100000, 150000, 150000, 230000, 134000, 456000, 321000, 123000, 123000), 
                      career_track = c(rep("technical", 4), rep("non-technical", 4), rep("executive", 2) ))

str(df_in_R)
summary(df_in_R)

```

Group by and aggregate functions using `dplyr`:

```{r}

df_in_R %>% 
  mutate(interval = cut(x = yrs_of_experience, 
                        breaks = seq(0, 30, 5), 
                        right = TRUE)) %>% 
  group_by(interval) %>% 
  summarize(count = n(), 
            mean = mean(compensation), 
            median = median(compensation), 
            min = min(compensation), 
            max = max(compensation), 
            std = sd(compensation)) %>% 
  ungroup()

```
# R verbs in Python

Dplyr verbs in Python using the `siuba` package.

Libraries:

```{python message = FALSE, eval = FALSE}
# siuba Package:
from siuba import _, group_by, summarize
```

Importing data from R in Python:

```{python eval = FALSE}
data = pd.DataFrame(df)

data.info()
data.describe()

```

```{python eval = FALSE}
(data >> 
  group_by(_.interval) >> 
  summarize(mean   = np.mean(_.compensation),
            median = (_.compensation).median(), # np.median()
            min = np.min(_.compensation), 
            max = np.max(_.compensation), 
            std = np.std(_.compensation))
  )

```




