---
title: "61 - Joining to Get Total Sales, a SQL Problem"
author: "Felipe Leandro Aguazaco"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: darkly
    highlight: zenburn
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      cache = TRUE,
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      dev = "png")
```

# Problem

---

Suppose you have two tables, one called 'orders', and one called 'revenue', as shown below: 

* Orders table:
```{r echo = FALSE}

library(tidyverse)
library(gt)
library(kableExtra)


orders <- data.frame(
  orders_id = seq(1, 6), 
  channel = c("Online", "Online", "In_Store", "In_Store", "Online", "Online"), 
  date = c("2018-09-01", "2018-09-03", "2018-10-11", "2018-08-21", "2018-08-13", "2018-10-29"), 
  month = c("September", "September", "October", "August", "August", "October")
)


#gt(orders) %>% 
  # tab_header(title = html("Tabla Información General")) %>% 
#  tab_style(style = list(cell_text(size = "medium", 
#                                   align = "center", 
#                                   v_align = "middle")),
#                         #cell_fill(color = "gray", alpha = 0.8)), 
#            locations = cells_body(columns = everything()))

orders %>% 
  kbl(align = "c", 
      col.names = c("Order Id", "Channel", "Date",  "Month")) %>%
  # kable_material(c("striped", "hover"), )
  kable_styling(full_width = FALSE, 
                bootstrap_options = c("striped", "hover"))
  # scroll_box(width = "100%", height = "200px")

```
* Revenue table:
```{r echo = FALSE}

revenue <- data.frame(order_id = seq(1, 6), 
                      revenue = c(100, 125, 200, 80, 200, 100))

revenue %>% 
  kbl(align = "c", 
      col.names = c("Order Id", "Revenue")) %>%
  # kable_material(c("striped", "hover"), )
  kable_styling(full_width = FALSE, 
                bootstrap_options = c("striped", "hover"))
  # scroll_box(width = "100%", height = "200px")

```
<!--
| order_id | channel  | date       | month     |
|----------|----------|------------|-----------|
| 1        | Online   | 2018-09-01 | September |
| 2        | Online   | 2018-09-03 | September |
| 3        | In_store | 2018-10-11 | October   |
| 4        | In_store | 2018-08-21 | August    |
| 5        | Online   | 2018-08-13 | August    |
| 6        | Online   | 2018-10-29 | October   |

* Revenue table:

| order_id | revenue |
|----------|---------|
| 1        | 100     |
| 2        | 125     |
| 3        | 200     |
| 4        | 80      |
| 5        | 200     |
| 6        | 100     |

-->

Using SQL, write a query to show the total revenue by channel for the months of September and October.

# DB in PostgreSQL 

---

Connecting a database in PostgreSQL to R.  

1. Libraries:

```{r}

library(DBI)
library(RPostgres)

```

2. Connection:

```{r}

question_61_db <- dbConnect(drv = Postgres(), 
                            dbname = "InterviewQs", 
                            host = "localhost", 
                            user = "postgres", 
                            password = Sys.getenv("PostgreSQL_Password"), 
                            bigint = "integer")

```

3. Connecting test:

```{r}
# Set schema
dbGetQuery(conn = question_61_db, 
           "SET SEARCH_PATH = question_61")

# Tables in data base
dbListTables(conn = question_61_db)

# Fields in a table
dbListFields(conn = question_61_db, 
             name = "orders")

# A simple query
dbGetQuery(conn = question_61_db, 
           "SELECT *
            FROM orders;")

```
# Solving the Problem {.tabset .tabset-fade .tabset-pills}

---

## In SQL

```{sql connection = question_61_db}
SELECT channel, 
       month,
       SUM(revenue) AS total_revunue
FROM orders
INNER JOIN revenue
  USING(order_id)
WHERE month !~ 'August'
GROUP BY channel, month
ORDER BY channel, month;

```
## In R

```{r}

library(dtplyr) # Database manipulation by data.table engine
library(tidyverse)

```

```{r}

answer_ques61 <- 
tbl(src = question_61_db, 
    "orders") %>% 
  lazy_dt() %>% 
  inner_join(y = tbl(src = question_61_db, 
                     "revenue") %>% 
               lazy_dt(), 
             by = "order_id") %>% 
  filter(month != "August") %>% 
  group_by(channel, month) %>% 
  summarize(total_revenue = sum(revenue)) %>% 
  ungroup() %>% 
  arrange(month, channel) %>% 
  collect()

```

```{r echo = FALSE}

library(kableExtra)

answer_ques61 %>% 
  kbl(align = "c", 
      col.names = c("Channel", "Month", "Total Revenue")) %>%
  # kable_material(c("striped", "hover"), )
  kable_styling(full_width = FALSE, 
                bootstrap_options = c("striped", "hover"))
  # scroll_box(width = "100%", height = "200px")

```

## In Python

```{r echo = FALSE}

library(reticulate)

orders <- tbl(src = question_61_db, 
              "orders") %>% 
  collect()

revenue <- tbl(src = question_61_db, 
              "revenue") %>% 
  collect()

```

Data:

```{python}
import pandas as pd

# Importing R object to Python
orders_py = r.orders
revenue_py = r.revenue

answer_ques61_py = pd.merge(left = orders_py, 
                            right = revenue_py, 
                            how = "inner", 
                            on = "order_id")

```

```{python}
answer_py = answer_ques61_py \
              .query("month != 'August'") \
              .groupby(by = ['channel', 'month']) \
              .agg(total_revenue = ('revenue', 'sum'))

print(answer_py)

```

