---
title: "64 - Sales by Marketing Channel"
author: "Felipe Leandro Aguazaco"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: darkly
    highlight: zenburn
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      #cache = TRUE,
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      #fig.width = 4, 
                      #fig.height = 4,
                      dev = "png")
```

# Problem

---

Given the following [data set](https://docs.google.com/spreadsheets/d/1s_gYPvxtgO9ufb-q2TO4n9_TCe7I-RorkbDYYegMm4c/edit#gid=33674467), can you plot a chart that shows the percent of revenue by marketing source? You can assume that purchase_value is revenue in this case, and you'll have to calculate the percent of revenue by source prior to plotting.

# Data

---

Libraries: 

```{r }

library(tidyverse)
library(dtplyr)
library(gsheet) # Read Google sheets
library(DataExplorer)
library(skimr)
library(viridis)
library(plotly)
library(htmltools)
library(reticulate)

```
Load data:
```{r}

data <- gsheet2tbl(url = "https://docs.google.com/spreadsheets/d/1s_gYPvxtgO9ufb-q2TO4n9_TCe7I-RorkbDYYegMm4c/edit#gid=33674467") %>% 
  tibble()

# Original data structure
str(data)

# Data Cleaning
data <- data %>% 
  mutate(across(.cols = c(2:3), 
                .fns = as.POSIXct)) %>% 
  mutate(across(.cols = c(6:8, 11), 
                .fns = as.factor))

```

# Data Exploratory Analysis

---

Summary:

```{r}

skim(data = data)

```

Graphs:

```{r}

# Final structure
str(data)

# Types of variables
data %>% 
  plot_intro(ggtheme = theme_bw())

# Percentage of missing values by variable
data %>% 
  plot_missing(ggtheme = theme_bw(), 
               theme_config = theme(legend.position = "none"))

# Distribution: histograms
data %>% 
  select_if(is.numeric) %>% # Not necessary
  select(-c(ip_address, user_id)) %>% 
  plot_histogram(geom_histogram_args = list(color = "black"), 
                 title = "Distribution Quantitative Variables", 
                 ggtheme = theme_bw(), 
                 ncol = 2,
                 theme_config = theme(plot.title = element_text(hjust = 0.5, 
                                                                face = "bold"), 
                                      axis.title.x = element_text(face = "plain"),
                                      axis.title.y = element_text(face = "plain")))

# Frecuency: barplots
data %>% 
  plot_bar(ggtheme = theme_bw(), 
           theme_config = theme(axis.text.x = element_text(size = 6.5)))

```

# Resolving in R

---

```{r}

div(
  ggplotly(
    data %>% 
    group_by(source) %>% 
    summarize(revenue = sum(purchase_value)) %>% 
    mutate(rev_percentage = round(revenue / sum(revenue) * 100, 
                                  digits = 2)) %>% 
    mutate(source = reorder(x = source, 
                            X = rev_percentage, 
                            FUN = min)) %>%
    ungroup() %>% 
    ggplot(mapping = aes(x = source, 
                         y = rev_percentage)) + 
      geom_col(width = 0.65, 
               colour = "black", 
               fill =  viridis::viridis(n = 3, 
                                        option = "E")) +
      labs(title = "Percent of Revenue by Marketing Source", 
           x = "Marketing Source", 
           y = "Percent of Revenue") + 
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5))
  ), 
align = "center")
  
```
# Resolving in Python

---

Importing data from R object to Python:

```{python}
data_python = r.data

data_python.dtypes

```

Aggregate data:

```{python}
import pandas as pd

# Problem
data_agg = data_python \
  .groupby(by = 'source', 
           as_index = False, 
           sort = False) \
  .agg(total_revenue = ('purchase_value', 'sum'))

data_agg['percent_revenue'] = round(data_agg.total_revenue / sum(data_agg.total_revenue) * 100, 2)

data_agg.dtypes

print(data_agg)

```

Graph:

```{python}
import plotly.express as px

fig = px.bar(data_agg, 
             x = 'source', 
             y = 'percent_revenue', 
             #height = 100, 
             #width = 100, 
             title = 'Percent of Revenue by Marketing Source',
             labels = {'source': 'Marketing Source',
             'percent_revenue': 'Percent of Revenue'}, 
             color = ["red", "gray", "blue"],
             color_discrete_map = "identity")

fig.show()

```




