---
title: "Data Science Interview Questions"
subtitle: "054 - Smallest Unrepresented Integer"
author: "Felipe Leandro Aguazaco"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: darkly
    highlight: zenburn
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      dev = "png")
```

* https://www.linkedin.com/in/felipe-leandro-aguazaco/  
* https://github.com/LeandroAguazaco  
* leandro.aguazaco@outlook.com  

# Problem 

Given a sorted (in increasing order) array of positive numbers, can you find the smallest positive integer that cannot be represented as a sum of elements from the array?

# Resolving Problem With R

Libraries:

```{r}

library(tidyverse) # Includes purrr package
library(gtools)
library(reticulate)

```

Sorted (in increasing order) array of positive numbers `vector`:

```{r}

vector <- c(1, 3, 6, 10, 11, 15)

```

Combinations in R using the `gtools::combinations` package and function:

```{r}

combinations(n = length(vector), # Amount of elements
             r = 3, # Selections 
             v = vector, # Array
             set = FALSE) 

```

Function that figure out the sum of all **combinations**, selecting n elements from a vector of positive integers:

```{r}

sum_combinations <- function(vector, n) {

  sums <- combinations(n = length(vector), 
                       r = n, 
                       v = vector, 
                       set = FALSE) %>% 
    addmargins(2) %>% # Sum rows
    data.frame() %>% 
    select(Sum)
  
  return(sums)

}

```

Applying the created function to obtain all possible numbers that can be represented as a sum of elements from the array:

```{r}

represented_integers <- map_dfr(.x = 1:length(vector), 
                                .f = ~ sum_combinations(vector, 
                                                        n = .x)) %>% 
  arrange(Sum) %>% 
  distinct() %>%  # Delete repeated elements
  pull(Sum)
  
# Integers that can be represented as a sum of elements from the vector
represented_integers

```

To obtain the minimum integer that can't be represented:

```{r}

# Integers that not appears in the sequence from 1 to maximum of represented numbers.
minimum_table <- data.frame(all_numbers = seq(1, max(represented_integers))) %>% 
  filter(!(all_numbers %in% represented_integers)) %>% 
  slice_min(order_by = all_numbers)

# Conditional selection of minimum integer if the number is between the range of represented integers or not.
if(dim(minimum_table)[1] == 0) {
  minimum_int <- max(represented_integers) + 1
} else {
  minimum_int <- minimum_table[1, 1]
}

```

The smallest positive integer that cannot be represented as a sum of elements from the array [`r vector`] is **`r minimum_int`**.

## Unique Function in R

Minimum unrepresented integer function:

```{r }

min_unrepresented_int <- function(vector) {
# 1.
  sum_combinations <- function(n) {
    sums <- combinations(n = length(vector), 
                         r = n, 
                         v = vector, 
                         set = FALSE) %>% 
      addmargins(2) %>% # Sum columns
      data.frame() %>% 
      select(Sum)
    
    return(sums)
  }
  
# 2. 
  represented_integers <- map_dfr(.x = 1:length(vector), 
                                  .f = ~ sum_combinations(n = .x)) %>% 
  arrange(Sum) %>% 
  distinct() %>%  # Delete repeated elements
  pull(Sum)
  
# 3. 
  minimum_table <- data.frame(all_numbers = seq(1, max(represented_integers), 1)) %>% 
    filter(!(all_numbers %in% represented_integers)) %>% 
    slice_min(order_by = all_numbers)
  
# 4. 
  if(dim(minimum_table)[1] == 0) {
    minimum_int <- max(represented_integers) + 1
  } else {
    minimum_int <- minimum_table[1, 1]
  }
  
# 5. 
  return(minimum_int)
}

```

Applying the unique R function:

```{r}

vector <- c(1, 3, 6, 10, 11, 15)

min_unrepresented_int(vector)

```

The smallest positive integer that cannot be represented as a sum of elements from the array [`r vector`] is **`r min_unrepresented_int(vector)`**.

```{r}

vector <- c(1, 1, 1, 1)

min_unrepresented_int(vector)

```

The smallest positive integer that cannot be represented as a sum of elements from the array [`r vector`] is **`r min_unrepresented_int(vector)`**.

```{r}

vector <- c(1, 1, 3, 4)

min_unrepresented_int(vector)

```

The smallest positive integer that cannot be represented as a sum of elements from the array [`r vector`] is **`r min_unrepresented_int(vector)`**.


# Resolving Problem With Python

Libraries:
<!-- class.source = "bg-info" # change chunck background-->
```{python}
import pandas as pd
import numpy as np
import itertools

```

Arrays in Python:

```{python}
# Array
array = [1, 3, 6, 10, 11, 15]

# Length of array
len(array)

```

Combinations in Python:

```{python}
combinations = pd.DataFrame(itertools.combinations(array, 3), 
                            columns = ["one", "two", "three"]) # Columns names

print(combinations)

# Sum rows:
combinations["sum"] = combinations.sum(1)
# combinations["sum"] = combinations.one + combinations.two + combinations.three

```

Function that calculates the sum of all possible combinations selecting *n* elements from an array:

```{python}
def sum_combinations(array, n):
  
  combinations = pd.DataFrame(list(itertools.combinations(array, 3))).sum(1)
  
  return combinations.sum(1)
  
```

Using `map` function to apply a function to an iterable object like arrays:

```{python}
represented_int = list(map(lambda n:
  pd.DataFrame(itertools.combinations(array, n)).sum(1), # Combinations Sum
  list(np.arange(1, len(array) + 1)))) # Iterable object, arguments for the lambda function
# Return a list of series

represented_int = pd.concat(represented_int, 0) # Concatenate Series by rows

represented_int = represented_int.drop_duplicates().sort_values() # Drop duplicates and sort elements

# represented_int.index = np.arange(0, len(represented_int), 1) # Replace indexes of a serie

type(represented_int) # Type of object
# dir(represented_int) # Attributes of the given object

```

To obtain the minimum integer that can't be represented:

```{python}
all_numbers = np.arange(1, represented_int.max() + 1)

# All elements that are in all_numbers but not represented_int
diff_int = list(set(all_numbers).difference(set(represented_int)))

# Conditional selection

if len(diff_int) != 0:
  minimum_int = min(diff_int)
else:
  minimum_int = represented_int.max() + 1

print(minimum_int)

```

## Unique Function in Python

```{python}
def min_unrepresented_int(array):
  # 1. 
  represented_int = list(map(lambda n:
    pd.DataFrame(itertools.combinations(array, n)).sum(1),
    list(np.arange(1, len(array) + 1)))) 
  
  represented_int = pd.concat(represented_int, 0)
  
  represented_int = represented_int.drop_duplicates().sort_values()
  
  # 2. 
  all_numbers = np.arange(1, represented_int.max() + 1)
  
  diff_int = list(set(all_numbers).difference(set(represented_int)))
  
  # 3.
  if len(diff_int) != 0:
    minimum_int = min(diff_int)
  else:
    minimum_int = represented_int.max() + 1 
  
  # 4.
  return minimum_int
```

Applying the unique Python function:

```{python}
integers = [1, 3, 6, 10, 11, 15]

min_unrepresented_int(integers)

```

The smallest positive integer that cannot be represented as a sum of elements from the array [`r py$integers`] is **`r py$min_unrepresented_int(py$integers)`**.

```{python}
integers = [1, 1, 1, 1]

min_unrepresented_int(integers)

```

The smallest positive integer that cannot be represented as a sum of elements from the array [`r py$integers`] is **`r py$min_unrepresented_int(py$integers)`**.

```{python}
integers = [1, 1, 3, 4]

min_unrepresented_int(integers)

```

The smallest positive integer that cannot be represented as a sum of elements from the array [`r py$integers`] is **`r py$min_unrepresented_int(py$integers)`**.

```{python}
integers = [15, 30, 45, 50]

min_unrepresented_int(integers)

```

The smallest positive integer that cannot be represented as a sum of elements from the array [`r py$integers`] is **`r py$min_unrepresented_int(py$integers)`**.